Windowed least-squares model reduction for dynamical systems

### 2019-10-17 section4/spacetime_numerical.tex
latest
small changes

\subsection{\spaceTimeAcronym\ trial spaces: direct and indirect methods}
We now briefly consider \spaceTimeAcronym\ trial subspaces. 
For \spaceTimeAcronym\ trial subspaces, the distinction between a direct and indirect method is less clear as the optimization variable (i.e., the generalized coordinates) are finite 
dimensional. There is no need to transcribe an infinite dimensional optimization variable into a finite dimensional one. There is, however, still a need to develop a finite 
dimensional representation of the objective function~\eqref{eq:obj_gen_slab}. In what follows, we describe two techniques to do this: one that works with the FOM O$\Delta$E and one that works with the FOM ODE. We associate direct methods as those 
that work with the O$\Delta$E and indirect methods with those that work directly with the ODE.

\subsection{\spaceTimeAcronym: direct solution approach}
The direct solution technique seeks to minimize the fully discrete objective function associated with the O$\Delta$E. For brevity, we again focus on linear multistep methods. For notational simplicity, we define an index mapping function that is 
equivalent to the mapping function $\indexMapper$, but outputs only the first argument: 
$$\indexMappern: (n,i) \mapsto 
\begin{cases}
1 & n=1, \; i = 0,\\
n & n \ge 1, \; i > 0, \\
\indexMappern(n-1,\nstepsArg{n-1}+i) & n > 1, \; i \le 0.
\end{cases}$$
%The definition of the fully discrete objective functions~\eqref{eq:obj_lms} as outlined in Section~\ref{sec:direct}, can be leveraged for this purpose. For simplicity, we only 
%outline the case for linear multistep schemes. 
Assuming $\text{Rank}(\stweightingMatOneArg{n})\nstepsArg{n} \ge \stdimArg{n}$, \methodAcronym\ leveraging an \spaceTimeAcronym\ basis sequentially computes the generalized coordinates $\stgenstateArg{n}$, $n=1,\ldots,\nslabs$, that satisfy
 \begin{equation}\label{eq:obj_gen_lms_final_st}
\begin{split}
& \underset{\stgenstatey}{\text{minimize } }
\objectiveArgLMS{n}_{\text{D-ST}} (\stbasisArg{n}{\timeWindowArg{n}{1}} \stgenstatey + \stateInterceptSTArg{n},\ldots,\stbasisArg{n}{\timeWindowArg{n}{\nstepsArg{n}}} \stgenstatey + \stateInterceptSTArg{n}),%  \\
%&\text{subject to } \;  \stbasisArg{n}{\timeStartArg{n}}\stgenstateyArg{n} = \begin{cases} 
%\mathbb{P}^n(\stbasisArgnt{n-1}\stgenstateArg{n-1})(\timeEndArg{n-1}) & n = 2,\ldots,\nslabs, \\ 
%\mathbb{P}^n(\stateFOMArgnt{1})(0) &
%n=1. \end{cases}
\end{split} 
\end{equation}
where the space--time objective function for the direct method is
$$
\objectiveArgLMS{n}_{\text{D-ST}}  :  (\stateyDiscreteArgnt{1},\ldots,\stateyDiscreteArgnt{\nstepsArg{n}}) \mapsto \objectiveArgLMS{n}_{\text{D}}(\stateyDiscreteArgnt{1},\ldots,\stateyDiscreteArgnt{\nstepsArg{n}}; \approxstateArg{\indexMappern(n,0)}{\timeWindow^{\indexMapper(n,0)}},\ldots, 
 \approxstateArg{\indexMappern(n,-\lmsWidthArg{n}{1}+1)}{\timeWindow^{\indexMapper(n,-\lmsWidthArg{n}{1}+1)}}).
$$
It is noted that the boundary conditions are automatically satisfied through the definition of the \spaceTimeAcronym\ trial subspace. The minimization problem~\eqref{eq:obj_gen_lms_final_st} again yields a least-squares problem.
\begin{remark}
For the limiting case where one window comprises entire domain (i.e., $\DeltaSlabArg{1}\equiv T$), uniform quadrature weights are used, the trial space is set to be $\stspaceSTArg{1} = \stspaceST$, the weighting matrices $\stweightingMatArg{1} = \lspgWeightingST = \mathbf{I}$, and $\nstepsArg{1} = N_t$ time instances are employed that satisfy $\timeWindowArg{1}{i} = t^i$, $i=1,\ldots,N_t$, then $\stgenstateArg{1} =  \stgenstate_\text{ST-LSPG}$ and direct \methodAcronym\ with an \spaceTimeAcronym\ trial subspace recovers ST-LSPG. 
\end{remark}

@@

 & \underset{\stgenstatey}{\text{minimize }} \mathcal{J}^n \bigg( \stbasisArgnt{n} \stgenstatey + \stateInterceptSTArg{n} \otimes \onesFunctionArg{n} \bigg) .%\\  

@@

discretization of the integral. Towards this end we introduce $\ncollocSTArg{n} \ge \text{ceil}(\stdimArg{n}/ \text{Rank}(\stweightingMatOneArg{n}))$ collocation points over the $n$th window, $ \{ \collocPointSTArg{n}{i} \}_{i=1}^{\ncollocSTArg{n}} \subseteq [\timeStartArg{n},\timeEndArg{n}]$, $n=1,\ldots,\nslabs$. 

@@

Leveraging these collocation points, the indirect method for \methodAcronym\ with an \spaceTimeAcronym\ trial subspace computes the generalized coordinates 

@@

$$\objectiveDiscreteSTArg{n}: \statey \mapsto \sum_{i=1}^{\ncollocSTArg{n}} \gamma_i 

@@

\bigg[ \dot{\statey}(\collocPointSTArg{n}{i})  - \velocity (\stateyArg{}{\collocPointSTArg{n}{i}},\collocPointSTArg{n}{i} ) \bigg].

@@

>>>>

\subsection{\spaceTimeAcronym\ trial spaces: direct and indirect methods}
We now briefly consider \spaceTimeAcronym\ trial subspaces. 
For \spaceTimeAcronym\ trial subspaces, the distinction between a direct and indirect method is less clear as the optimization variable (i.e., the generalized coordinates) are finite 
dimensional. There is no need to transcribe an infinite dimensional optimization variable into a finite dimensional one. There is, however, still a need to develop a finite 
dimensional representation of the objective functional~\eqref{eq:obj_gen_slab}. In what follows, we describe two techniques to do this: one that works with the FOM O$\Delta$E and one that works with the FOM ODE. We associate direct methods as those 
that work with the O$\Delta$E and indirect methods with those that work directly with the ODE.

\subsection{\spaceTimeAcronym: direct solution approach}
The direct solution technique seeks to minimize the fully discrete objective function associated with the O$\Delta$E. For brevity, we again focus on linear multistep methods and leverage the time grids introduced in Section~\ref{sec:direct}. For notational simplicity, we define an index mapping function that is 
equivalent to the mapping function $\indexMapper$, but outputs only the first argument: 
$$\indexMappern: (n,i) \mapsto 
\begin{cases}
1 & n=1, \; i = 0,\\
n & n \ge 1, \; i > 0, \\
\indexMappern(n-1,\nstepsArg{n-1}+i) & n > 1, \; i \le 0.
\end{cases}$$
%The definition of the fully discrete objective functions~\eqref{eq:obj_lms} as outlined in Section~\ref{sec:direct}, can be leveraged for this purpose. For simplicity, we only 
%outline the case for linear multistep schemes. 
Assuming $\text{Rank}(\stweightingMatOneArg{n})\nstepsArg{n} \ge \stdimArg{n}$, \methodAcronym\ with the direct approach and an \spaceTimeAcronym\ trial subspace sequentially computes the generalized coordinates $\stgenstateArg{n}$, $n=1,\ldots,\nslabs$ that satisfy
 \begin{equation}\label{eq:obj_gen_lms_final_st}
\begin{split}
& \underset{\stgenstatey \in \RR{\stdimArg{n}}}{\text{minimize } }
\objectiveArgLMS{n}_{\text{D-ST}} (\stbasisArg{n}{\timeWindowArg{n}{1}} \stgenstatey + \stateInterceptSTArg{n},\ldots,\stbasisArg{n}{\timeWindowArg{n}{\nstepsArg{n}}} \stgenstatey + \stateInterceptSTArg{n}),%  \\
%&\text{subject to } \;  \stbasisArg{n}{\timeStartArg{n}}\stgenstateyArg{n} = \begin{cases} 
%\mathbb{P}^n(\stbasisArgnt{n-1}\stgenstateArg{n-1})(\timeEndArg{n-1}) & n = 2,\ldots,\nslabs, \\ 
%\mathbb{P}^n(\stateFOMArgnt{1})(0) &
%n=1. \end{cases}
\end{split} 
\end{equation}
where the space--time objective function for the direct method is
$$
\objectiveArgLMS{n}_{\text{D-ST}}  :  (\stateyDiscreteArgnt{1},\ldots,\stateyDiscreteArgnt{\nstepsArg{n}}) \mapsto \objectiveArgLMS{n}_{\text{D}}(\stateyDiscreteArgnt{1},\ldots,\stateyDiscreteArgnt{\nstepsArg{n}}; \approxstateArg{\indexMappern(n,0)}{\timeWindow^{\indexMapper(n,0)}},\ldots, 
 \approxstateArg{\indexMappern(n,-\lmsWidthArg{n}{1}+1)}{\timeWindow^{\indexMapper(n,-\lmsWidthArg{n}{1}+1)}}).
$$
It is noted that the boundary conditions are automatically satisfied through the definition of the \spaceTimeAcronym\ trial subspace. The minimization problem~\eqref{eq:obj_gen_lms_final_st} again yields a least-squares problem.
\begin{remark}
For the limiting case where one window comprises entire domain (i.e., $\DeltaSlabArg{1}\equiv T$), uniform quadrature weights are used, the trial space is set to be $\stspaceSTArg{1} = \stspaceST$, the weighting matrices $\stweightingMatArg{1} = \lspgWeightingST = \mathbf{I}$, and $\nstepsArg{1} = N_t$ time instances are employed that satisfy $\timeWindowArg{1}{i} = t^i$, $i=1,\ldots,N_t$, then $\stgenstateArg{1} =  \stgenstate_\text{ST-LSPG}$ and direct \methodAcronym\ with an \spaceTimeAcronym\ trial subspace recovers ST-LSPG. 
\end{remark}

@@

 & \underset{\stgenstatey \in \RR{\stdimArg{n}}}{\text{minimize }} \mathcal{J}^n \bigg( \stbasisArgnt{n} \stgenstatey + \stateInterceptSTArg{n} \otimes \onesFunctionArg{n} \bigg) .%\\  

@@

discretization of the integral. Towards this end we introduce $\ncollocSTArg{n} \ge \text{ceil}(\stdimArg{n}/ \text{Rank}(\stweightingMatOneArg{n}))$ quadrature points over the $n$th window, $ \{ \collocPointSTArg{n}{i} \}_{i=1}^{\ncollocSTArg{n}} \subseteq [\timeStartArg{n},\timeEndArg{n}]$, $n=1,\ldots,\nslabs$. 

@@

Leveraging these quadrature points, \methodAcronym\ with the indirect method and an \spaceTimeAcronym\ trial subspace computes the generalized coordinates 

@@

$$\objectiveDiscreteSTArg{n}: \statey \mapsto \sum_{i=1}^{\ncollocSTArg{n}} \zeta^{n,i} 

@@

\bigg[ \dot{\statey}(\collocPointSTArg{n}{i})  - \velocity (\stateyArg{}{\collocPointSTArg{n}{i}},\collocPointSTArg{n}{i} ) \bigg],

@@

where $\zeta^{n,i} \in \RRplus$, $i=1,\ldots,\ncollocSTArg{n}$ are quadrature weights over the $n$th window. The optimization problem~\eqref{eq:obj_gen_slab2} again comprises a least-squares problem.

### 2019-10-17 section4/spatial_numerical.tex
small changes

The FBSM is an iterative approach that can be used to
solve the two-point boundary value problem. The general process of the FBSM is
as follows: First, the system~\eqref{eq:lspg_continuous} is (numerically) solved forward in
time using an initial guess to the costate to obtain an initial guess to
the generalized coordinates. Next, the adjoint equation~\eqref{eq:lspg_adjoint} is solved \textit{backwards} in time given the
approximation to the generalized coordinates. This gives a new estimate to the
costate, which is then used to again solve the forward problem to obtain
a new estimate of the generalized coordinates. This process is continued until
convergence. Algorithm~\ref{alg:st_iter} outlines the method. The algorithm
contains three parameters: the damping factor $\rho \le 1$, the growth factor
$\fbsmGrowth \ge 1$, and the decay factor $\fbsmDecay \ge 1$. The damping factor controls the rate at which the costate seen by~\eqref{eq:lspg_continuous} is updated. The
closer $\rho$ is to unity, the faster the resulting algorithm will converge.
For large window sizes, however, too high a value of $\rho$ can lead to an unstable iterative process. 
A proper value of $\rho$ can be
obtained with a line search. The line search presented in Algorithm~\ref{alg:st_iter} adapts the damping factor
according to the objective. Convergence properties of the FBSM method are
presented in Ref.~\cite{McAsey2012ConvergenceOT}. It is shown that, for a small 
enough value of $\rho$, the algorithm will converge.

\begin{algorithm} \caption{\spatialAcronym\ trial subspace: algorithm for the FBSM over the $n$th window.} \label{alg:st_iter} 
\SetKwInOut{Input}{Input}\SetKwInOut{Output}{Output}
\Input{tolerance, $\epsilon$; damping factor, $\rho \le 1$; growth factor, $\fbsmGrowth \ge 1$; decay
factor, $\fbsmDecay \ge 1$} 
\Output{Stationary point, $\genstateArgnt{n}$ }
\textbf{Online Steps:}\\ 
$\genstate^n_0 \leftarrow \bz$ \Comment{Set initial guess for state} \\
$\controllerArgnt{n} \leftarrow \boldsymbol 0$ \Comment{Set initial
guess for costate}\\ 
$\text{Compute } \genstateArgnt{n}_1 \text{ satisfying }  \massArg{n} \genstateDotArgnt{n}_1(t)  -  \basisspaceTArg{n} \stweightingMatArg{n}
\velocity(\veloargsromArg{1}) =  \massArg{n} \controllerArg{n}{t}$ 
% Eq.~\ref{eq:lspg_continuous} with $\adjoint^n(t) = \controllerArg{n}{t}$ to
% obtain $\genstate_0^{n}(t)$ 
\Comment{Solve~\eqref{eq:lspg_continuous}}\\ 
%Set: $\genstate^n_1 = \genstate$ \Comment{Set state} \\
%$\objectiveArg{n}({\genstate_1^n(t)})$
%\Comment{Evaluate objective function}\\
$i \leftarrow 1$ \Comment{Set counter}\\
\While{$\epsilon \le \int_{\timeStartArg{n}}^{\timeEndArg{n}} \norm{\genstate^n_{i}(t) - \genstate^n_{i-1}(t) }dt $}{
\small{
\begin{multline*}
\text{ Compute } \adjointArgnt{n} \text{ satisfying }
\massArg{n} \adjointDotArg{n}{t}  + \basisspaceTArg{n} \bigg[\frac{\partial \velocity}{\partial \stateyDiscrete}(\basisspaceArg{n} \genstate^n_i(t) + \stateInterceptArg{n},t) \bigg]^T \stweightingMatArg{n} \basisspaceArg{n} \adjointArg{n}{t}= \\ -\bigg[\basisspaceTArg{n} \bigg[ \frac{\partial \velocity}{\partial \stateyDiscrete} ( \basisspaceArg{n} \genstate^n_i(t) + \stateInterceptArg{n},t) \bigg]^T \stweightingMatArg{n} \bigg( \mathbf{I} -   \basisspaceArg{n} [\massArg{n}]^{-1} \basisspaceTArg{n} \stweightingMatArg{n} \bigg)  \bigg( \basisspaceArg{n} \dot{\genstate}_i^n(t)   -   \velocity( \basisspaceArg{n} \genstate_i^n(t)  +\stateInterceptArg{n},t) \bigg) \bigg] 
\end{multline*} }
\Comment{Solve~\eqref{eq:lspg_adjoint} to obtain guess to costate} \\
%Solve Eq.~\ref{eq:lspg_adjoint} with $\genstate^n(t) = \genstate^n_i(t)$ to obtain $\adjoint^n(t)$ \\
$\controllerArgnt{n}  \leftarrow \rho \controllerArgnt{n} + (1 - \rho) \adjointArgnt{n}$ \Comment{Weighted update to costate}\\
$i \leftarrow i+1$ \Comment{Update counter}\\
$\text{Compute }\genstateArgnt{n}_i \text{ satisfying } \massArg{n} \genstateDotArgnt{n}_i(t)   -  \basisspaceTArg{n} \stweightingMatArg{n} \velocity(\basisspaceArg{n} \genstate^n_i(t) + \stateInterceptArg{n},t) = \massArg{n} \controllerArg{n}{t} $
\Comment{Solve~\eqref{eq:lspg_continuous}}\\
%$\objectiveArg{n}({\genstate_i^n(t)})$ \Comment{Evaluate objective function}\\
\uIf{ $\objectiveArg{n}({\basisspaceArg{n}\genstate_i^n + \stateInterceptArg{n}\otimes \onesFunctionArg{n}}) \le \objectiveArg{n}({\basisspaceArg{n}\genstate_{i-1}^n + \stateInterceptArg{n}\otimes \onesFunctionArg{n}})$}
{
$\rho \leftarrow \text{min}(\rho \fbsmGrowth,1)$ \Comment{Grow the damping factor}\\
}
\Else{
$\rho \leftarrow \frac{\rho }{ \fbsmDecay}$ \Comment{Shrink the damping factor}\\ 
$\genstate_i^{n} \leftarrow  \genstate_{i-1}^{n}$ \Comment{Reset state to value at previous iteration}
}
}
Return converged solution, $\genstateArgnt{n}= \genstateArgnt{n}_i$
\end{algorithm}
\subsubsection{Numerical considerations for the forward and backward problems}
The FBSM requires solving the systems~\eqref{eq:lspg_continuous} and~\eqref{eq:lspg_adjoint}, both of which are defined at the time-continuous level. 
The numerical implementation of the FBSM requires two main ingredients: (1) temporal discretization schemes for the forward and backward problems and (2) 
an efficient method for computing the action of the Jacobian transpose on a vector.

@@

The second ingredient, namely an efficient method for computing the action of the Jacobian transpose on a vector, can be challenging. 
In the case one does not have access to the full-order model Jacobian (or it is too costly to compute), the evaluation of this term can be both challenging and expensive. Here, two methods are discussed that can be used to evaluate the action of the Jacobian transpose on a vector as seen in~\eqref{eq:lspg_adjoint}:
\begin{enumerate}
\item \textit{Jacobian-free approximation}: A non-intrusive way to evaluate the Jacobian transpose in~\eqref{eq:lspg_adjoint} is to recognize that all terms including the Jacobian transpose are left multiplied by the transpose of the trial basis. One can make the manipulation,
$$\basisspaceTArg{n} \bigg[\frac{\partial \velocity}{\partial \stateyDiscrete} (\veloargsromn)\bigg]^T = \bigg[  \frac{\partial \velocity}{\partial \stateyDiscrete} (\veloargsromn) \basisspaceArg{n} \bigg]^T.$$
This manipulation allows one to compute the action of the Jacobian on each column of $\basisspaceArg{n}$ by, e.g., the finite difference approximation
$$\frac{\partial \velocity}{\partial \stateyDiscrete}(\veloargsromn) \basisvec_i^n \approx \frac{1}{\epsilon}\bigg( \velocity(\basisspaceArg{n}\genstateArg{n}{t} + \stateInterceptArg{n} + \epsilon \basisvec_i^n,t) - \velocity(\veloargsromn) \bigg).$$
The Jacobian transpose can be formed in $K+1$ evaluations of the velocity. For cases where the reduced-order model is low dimensional, this approach is feasible. For higher dimensional reduced-order models, this approach can be prohibitively expensive.

>>>>

The FBSM is an iterative approach that can be used to
solve the two-point boundary value problem. The general process of the FBSM is
as follows: First, the system~\eqref{eq:lspg_continuous} is (numerically) solved forward in
time using an initial guess to the costate to obtain an initial guess to
the generalized coordinates. Next, the adjoint equation~\eqref{eq:lspg_adjoint} is solved \textit{backwards} in time given the
approximation to the generalized coordinates. This gives a new estimate to the
costate, which is then used to again solve the forward problem to obtain
a new estimate of the generalized coordinates. This process is continued until
convergence. Algorithm~\ref{alg:st_iter} outlines the method. The algorithm
contains three parameters: the relaxation factor $\rho \le 1$, the growth factor
$\fbsmGrowth \ge 1$, and the decay factor $\fbsmDecay \ge 1$. The relaxation factor controls the rate at which the costate seen by~\eqref{eq:lspg_continuous} is updated. The
closer $\rho$ is to unity, the faster the algorithm will converge.
For large window sizes, however, too high a value of $\rho$ can lead to an unstable iterative process. 
A proper value of $\rho$ can be
obtained with a line search. The line search presented in Algorithm~\ref{alg:st_iter} adapts the relaxation factor
according to the objective. Convergence properties of the FBSM method are
presented in Ref.~\cite{McAsey2012ConvergenceOT}. It is shown that, for a small 
enough value of $\rho$, the algorithm will converge.

\begin{algorithm} \caption{\spatialAcronym\ trial subspace: algorithm for the FBSM over the $n$th window.} \label{alg:st_iter} 
\SetKwInOut{Input}{Input}\SetKwInOut{Output}{Output}
\Input{tolerance, $\epsilon$; relaxation factor, $\rho \le 1$; growth factor, $\fbsmGrowth \ge 1$; decay
factor, $\fbsmDecay \ge 1$} 
\Output{Stationary point, $\genstateArgnt{n}$ }
\textbf{Online Steps:}\\ 
$\genstate^n_0 \leftarrow \bz$ \Comment{Set initial guess for state} \\
$\controllerArgnt{n} \leftarrow \boldsymbol 0$ \Comment{Set initial
guess for costate}\\ 
$\text{Compute } \genstateArgnt{n}_1 \text{ satisfying }  \massArg{n} \genstateDotArgnt{n}_1(t)  -  \basisspaceTArg{n} \stweightingMatArg{n}
\velocity(\veloargsromArg{1}) =  \massArg{n} \controllerArg{n}{t}$ 
% Eq.~\ref{eq:lspg_continuous} with $\adjoint^n(t) = \controllerArg{n}{t}$ to
% obtain $\genstate_0^{n}(t)$ 
\Comment{Solve~\eqref{eq:lspg_continuous}}\\ 
%Set: $\genstate^n_1 = \genstate$ \Comment{Set state} \\
%$\objectiveArg{n}({\genstate_1^n(t)})$
%\Comment{Evaluate objective function}\\
$i \leftarrow 1$ \Comment{Set counter}\\
\While{$\epsilon \le \int_{\timeStartArg{n}}^{\timeEndArg{n}} \norm{\genstate^n_{i}(t) - \genstate^n_{i-1}(t) }dt $}{
\small{
\begin{multline*}
\text{ Compute } \adjointArgnt{n} \text{ satisfying }
\massArg{n} \adjointDotArg{n}{t}  + \basisspaceTArg{n} \bigg[\frac{\partial \velocity}{\partial \stateyDiscrete}(\basisspaceArg{n} \genstate^n_i(t) + \stateInterceptArg{n},t) \bigg]^T \stweightingMatArg{n} \basisspaceArg{n} \adjointArg{n}{t}= \\ -\bigg[\basisspaceTArg{n} \bigg[ \frac{\partial \velocity}{\partial \stateyDiscrete} ( \basisspaceArg{n} \genstate^n_i(t) + \stateInterceptArg{n},t) \bigg]^T \stweightingMatArg{n} \bigg( \mathbf{I} -   \basisspaceArg{n} [\massArg{n}]^{-1} \basisspaceTArg{n} \stweightingMatArg{n} \bigg)  \bigg( \basisspaceArg{n} \dot{\genstate}_i^n(t)   -   \velocity( \basisspaceArg{n} \genstate_i^n(t)  +\stateInterceptArg{n},t) \bigg) \bigg] 
\end{multline*} }
\Comment{Solve~\eqref{eq:lspg_adjoint} to obtain guess to costate} \\
%Solve Eq.~\ref{eq:lspg_adjoint} with $\genstate^n(t) = \genstate^n_i(t)$ to obtain $\adjoint^n(t)$ \\
$\controllerArgnt{n}  \leftarrow \rho \controllerArgnt{n} + (1 - \rho) \adjointArgnt{n}$ \Comment{Weighted update to costate}\\
$i \leftarrow i+1$ \Comment{Update counter}\\
$\text{Compute }\genstateArgnt{n}_i \text{ satisfying } \massArg{n} \genstateDotArgnt{n}_i(t)   -  \basisspaceTArg{n} \stweightingMatArg{n} \velocity(\basisspaceArg{n} \genstate^n_i(t) + \stateInterceptArg{n},t) = \massArg{n} \controllerArg{n}{t} $
\Comment{Solve~\eqref{eq:lspg_continuous}}\\
%$\objectiveArg{n}({\genstate_i^n(t)})$ \Comment{Evaluate objective function}\\
\uIf{ $\objectiveArg{n}({\basisspaceArg{n}\genstate_i^n + \stateInterceptArg{n}\otimes \onesFunctionArg{n}}) \le \objectiveArg{n}({\basisspaceArg{n}\genstate_{i-1}^n + \stateInterceptArg{n}\otimes \onesFunctionArg{n}})$}
{
$\rho \leftarrow \text{min}(\rho \fbsmGrowth,1)$ \Comment{Grow the relaxation factor}\\
}
\Else{
$\rho \leftarrow \frac{\rho }{ \fbsmDecay}$ \Comment{Shrink the relaxation factor}\\ 
$\genstate_i^{n} \leftarrow  \genstate_{i-1}^{n}$ \Comment{Reset state to value at previous iteration}
}
}
Return converged solution, $\genstateArgnt{n}= \genstateArgnt{n}_i$
\end{algorithm}
\subsubsection{Numerical considerations for the forward and backward problems}
The FBSM requires solving the systems~\eqref{eq:lspg_continuous} and~\eqref{eq:lspg_adjoint}, both of which are defined at the time-continuous level. 
The numerical implementation of the FBSM requires two main ingredients: (1) temporal discretization schemes for the forward and backward problems and (2) 
an efficient method for computing the action of the Jacobian transpose on a vector.

@@

The second ingredient, namely an efficient method for computing the action of the Jacobian transpose on a vector, can be challenging. 
In the case one does not have access to the full-order model Jacobian (or it is too costly to compute), the evaluation of this term can be challenging and expensive. Here, two methods are discussed that can be used to evaluate the action of the Jacobian transpose on a vector as seen in~\eqref{eq:lspg_adjoint}:
\begin{enumerate}
\item \textit{Jacobian-free approximation}: A non-intrusive way to evaluate the Jacobian transpose in~\eqref{eq:lspg_adjoint} is to recognize that all terms including the Jacobian transpose are left multiplied by the transpose of the trial basis. One can make the manipulation,
$$\basisspaceTArg{n} \bigg[\frac{\partial \velocity}{\partial \stateyDiscrete} (\veloargsromn)\bigg]^T = \bigg[  \frac{\partial \velocity}{\partial \stateyDiscrete} (\veloargsromn) \basisspaceArg{n} \bigg]^T.$$
This manipulation allows one to compute the action of the Jacobian on each column of $\basisspaceArg{n}$ by, e.g., the finite difference approximation
$$\frac{\partial \velocity}{\partial \stateyDiscrete}(\veloargsromn) \basisvec_i^n \approx \frac{1}{\epsilon}\bigg( \velocity(\basisspaceArg{n}\genstateArg{n}{t} + \stateInterceptArg{n} + \epsilon \basisvec_i^n,t) - \velocity(\veloargsromn) \bigg).$$
The Jacobian transpose can be formed in $\romdimArg{n}+1$ evaluations of the velocity. For cases where the reduced-order model is low dimensional, this approach is feasible. For higher dimensional reduced-order models, this approach can be prohibitively expensive.
